Decomposition-based Job-shop Scheduling with Constrained Clustering

### 2021-10-12 Artificial Intelligence and Machine Learning/PADL_conference/Paper_draft_v1.tex
went over introduction

However, JSP is an NP-complete combinatorial optimization problem \cite{DBLP:journals/mor/GareyJS76,SOTSKOV1995237} for which no efficient algorithms are known. Therefore, searching for an optimal solution with state-of-the-art solvers for Answer Set Programming (ASP), Mixed Integer Programming, or Constraint Programming \cite{meng2020mixed,coltep19a,el2020job,al2017job}
%\todo{cite at least 1 paper solving JSP with these methods} 
often takes too much computation time, even for seemingly small instances. Practical applications instead necessitate solving JSP instances of large scale with thousands of operations~\cite{zhang2010hybrid}.
As a result, much research work focuses on % the development of 
efficient methods for finding high-quality approximations of optimal schedules, including dispatching rules and other heuristic approaches~\cite{blackstone1982state} as well as stochastic optimization techniques \cite{DBLP:journals/informs/VaessensAL96,DBLP:journals/jim/CalisB15}.  
The main issue of these approaches is that they require manual parametrization for a particular scheduling problem, which is tedious and error-prone. For instance, adapting heuristic methods might involve the development of new dispatching rules or specific combinations of existing ones. Similarly, merely utilizing default parameters for stochastic techniques might lead to mediocre schedules.    
In view of these challenges, recent research interest lies on the automated parametrization of approximation methods using machine learning methodologies~\cite{bengio2020machine}.

@@

In this work, we suggest a clustering-based method that automatically decomposes a JSP instance into sub-instances which can be tackled by 
optimization algorithms of an ASP solver. 
General clustering algorithms are unsupervised learning methods that partition a given set of objects into disjoint clusters, where each cluster comprises closest objects wrt.\ some distance function. 
In scheduling settings, however, clusters must satisfy additional constraints, which define precedence relations between operations of a job. Standard constrained clustering algorithms are not applicable in scheduling scenarios since they usually allow for the definition of disjointness constraints, which specify objects that cannot appear in one cluster \cite{zhang2019framework,wagstaff2001constrained,ding2020unified}. 
Therefore, our clustering method implements a novel type of constraints that
\begin{enumerate*}[label=\emph{(\roman*)}]
  \item prevent the assignment of an operation to a cluster if its preceding operation is not yet assigned to the same or previous cluster, and
  \item ensure balancing between the clusters according to the number of operations per cluster.
\end{enumerate*}
Further contributions of this paper can be summarized as follows:
\begin{itemize}
  \item Since a typical JSP instance describing jobs, their operations, and available machines, does not provide enough information for a clustering algorithm to find a good decomposition, we suggest various heuristic feature extractors, such as First-In-First-Out or machine load.
  \item We implement the suggested constrained clustering method and combine it with the forward selection of features. The latter is an automatic procedure aiming to find a subset of features allowing the clustering algorithm to compute a decomposition resulting in the best quality schedule.
  \item Evaluation of our method on the Taillard benchmark instances \cite{taillard1993benchmarks} showed that it significantly outperforms baseline heuristic methods and pure ASP optimization algorithms wrt.\ the makespan optimization criteria within the timeout of 600 seconds.
\end{itemize}
%It can be said that the attributes of the classical JSP are quite a few and not sufficient to define the similarity and dissimilarity between the operations. The lack of these attributes guides us to extract more features that could specify which operations are similar and should be in the same window and which shall not be in the same window. The main contribution of this work is to extract some features that have a significant impact on the decomposition process and accordingly on the makespan. 
%More specifically, we need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time. \\

@@

The paper is organized as follows: the following section provides the definition of JSP and gives an illustrative example of input data. Section \ref{sec:features} discusses heuristic feature extractors, followed by Section \ref{sec:method}, which presents the proposed constrained clustering algorithm. Section \ref{sec:eval} presents an extensive computational study on benchmark instances, comparing our results with the baseline solvers. A literature review is given in Section \ref{sec:literature}, followed by the conclusions and future work.

\section{Problem Definition}
The Job-shop Scheduling Problem (JSP) is one of the well-known scheduling problems \cite{baker1974introduction,lenstra1979computational,taillard1993benchmarks}. An instance of a JSP comprises a set $J = \{ J_1, J_2, \dots ,J_n \}$  of jobs and a set $M = \{ M_1, M_2, \dots ,M_m \}$ of Machines. Each job $J_i$ has a sequence $(O_{i,1}, O_{i,2}, \dots ,O_{i,m})$ of operations that must be processed in a predefined order. Each machine $M_i$ executes an operation of each job with a known and fixed processing time. The main objective of a JSP solver is to find a schedule that minimizes some criteria like makespan, i.e., the total completion time for all jobs.
In addition, JSP makes the following assumptions:
\begin{itemize}
	\item Once an operation starts, it must be completed without interruptions.
	\item Each job has no duplicated operations, i.e., visits a machine only once.
	\item A machine cannot perform more than one operation at a time.
	\item All machines are available at all times.
	\item Processing the jobs starts at time 0.
\end{itemize}

>>>>

However, JSP is an NP-complete combinatorial optimization problem \cite{DBLP:journals/mor/GareyJS76,SOTSKOV1995237} for which no efficient algorithms are known. Therefore, searching for an optimal solution with state-of-the-art solvers for Answer Set Programming (ASP), Mixed Integer Programming, or Constraint Programming \cite{meng2020mixed,coltep19a,el2020job,al2017job}
%\todo{cite at least 1 paper solving JSP with these methods} 
often takes too much computation time, even for seemingly small instances. Practical applications instead necessitate solving JSP instances of large scale with thousands of operations~\cite{zhang2010hybrid}.
As a result, much research work focuses on % the development of 
efficient methods for finding high-quality approximations of optimal schedules, including dispatching rules and other heuristic approaches~\cite{blackstone1982state} as well as stochastic optimization techniques \cite{DBLP:journals/informs/VaessensAL96,DBLP:journals/jim/CalisB15}.  
The main issue of these approaches is that they require manual parametrization for a particular scheduling problem, which is tedious and error-prone. For instance, adapting heuristic methods might involve the development of new dispatching rules or specific combinations of existing ones. Similarly, merely utilizing default parameters for stochastic techniques might lead to mediocre schedules.    
In view of these challenges, recent research interest lies on the automatic parametrization of approximation methods using machine learning methodologies~\cite{bengio2020machine}.

@@

In this work, we introduce a method based on clustering to automatically decomposes a JSP instance into several time windows that are small enough for optimization by an ASP solver.
General clustering algorithms are unsupervised learning methods that partition a given set of objects into disjoint clusters, where each cluster comprises close objects wrt.\ some distance measure. 
In scheduling settings, however, clusters must satisfy additional constraints implied by the precedence relation between operations of a job. Standard constrained clustering algorithms are not readily applicable to scheduling scenarios either, as they are limited to disjointness constraints specifying objects that must not appear together in a cluster \cite{zhang2019framework,wagstaff2001constrained,ding2020unified}. 
Therefore, our clustering method implements a novel type of constraints that
\begin{enumerate*}[label=\emph{(\roman*)}]
  \item prevent the assignment of an operation to a cluster if its preceding operation is not yet assigned to the same or a previous cluster and
  \item ensure balancing between clusters according to the target number of operations per cluster.
\end{enumerate*}
Further contributions of our work can be summarized as follows:
\begin{itemize}
  \item Since a typical JSP instance describing jobs, their operations, and available machines does not provide sufficient information by itself for finding some promising decomposition by a clustering algorithm, we incorporate heuristic approaches like \textit{First-In-First-Out} and \textit{Machine Load} to extract features from their solutions.
  \item We implement the proposed constrained clustering algorithm and combine it with the forward selection of features, which is an automatic method for identifying a subset of features allowing the clustering method to compute decompositions resulting in best-quality schedules.
  \item The evaluation of our approach on Taillard's benchmark instances \cite{taillard1993benchmarks} shows that it significantly outperforms baseline heuristic methods and pure ASP optimization wrt.\ the makespan optimization criterion within a timeout limit of 600 seconds.
\end{itemize}
%It can be said that the attributes of the classical JSP are quite a few and not sufficient to define the similarity and dissimilarity between the operations. The lack of these attributes guides us to extract more features that could specify which operations are similar and should be in the same window and which shall not be in the same window. The main contribution of this work is to extract some features that have a significant impact on the decomposition process and accordingly on the makespan. 
%More specifically, we need to extract some features to define which operations are similar to be scheduled in the same window and get a near-optimal solution in a reasonable time. \\

@@

The rest of this paper is organized as follows. Section~\ref{sec:problem} introduces JSP along with a running example. In Section \ref{sec:features}, we describe the feature extraction, including heuristic methods to obtain corresponding reference solutions. Section~\ref{sec:method} presents our proposed constrained clustering algorithm. In Section~\ref{sec:eval}, we empirically evaluate our approach and compare it to baseline heuristic methods as well as pure ASP optimization. Section~\ref{sec:literature} then surveys related work, followed by conclusions and future work in Section~\ref{sec:conclusions}.

\section{Problem Definition}\label{sec:problem}
The Job-shop Scheduling Problem (JSP) is one of the well-known scheduling problems \cite{baker1974introduction,lenstra1979computational,taillard1993benchmarks}. An instance of a JSP comprises a set $J = \{ J_1, J_2, \dots ,J_n \}$  of jobs and a set $M = \{ M_1, M_2, \dots ,M_m \}$ of Machines. Each job $J_i$ has a sequence $(O_{i,1}, O_{i,2}, \dots ,O_{i,m})$ of operations that must be processed in a predefined order. Each machine $M_i$ executes an operation of each job with a known and fixed processing time. The main objective of a JSP solver is to find a schedule that minimizes some criteria like makespan, i.e., the total completion time for all jobs.
In addition, JSP makes the following assumptions:
\begin{itemize}
	\item Once an operation starts, it must be completed without interruptions.
	\item Each job has no duplicated operations, i.e., visits a machine only once.
	\item A machine cannot perform more than one operation at a time.
	\item All machines are available at all times.
	\item Processing the jobs starts at time 0.
\end{itemize}
